Time Stretching of the GeV Emission of GRBs: Fermi LAT data vs geometrical model \\ {\large \url{https://github.com/maxitg/GammaRays}}

### 2014-08-28 Paper/gammaRays.tex
Replaced many links to Tools with two links

Before reading the data, we need to apply the basic filtering as described in Fermi guidelines (the actual code, and all parameters values we use to invoke Fermi Science tools are listed in appendix \ref{sec:fermiCode}).
Two tools from the Fermi Science Tools package are used for that task (the same filtering is applied to both burst and background data).

First, we use \texttt{gtselect}\footnote{\url{http://fermi.gsfc.nasa.gov/ssc/data/analysis/scitools/help/gtselect.txt}}.
We filter the Earth limb emission with it below the zenith angle of $100\degree$.
We also filter out photons with transient event class for a reason described in the next section.

The second one we use is \texttt{gtmktime}\footnote{\url{http://fermi.gsfc.nasa.gov/ssc/data/analysis/scitools/help/gtmktime.txt}}. It uses the spacecraft file as well as events file, and filters out time periods when Fermi LAT was not operational. Running \texttt{gtmktime} is also required for exposure maps calculation, which we'll need later on.

Now, with basic filtering done, we can read events data from \texttt{timed.fits}. This file, however, contains observations of a large area of the sky.
Before using the data obtained from it, we should perform a more elaborate filtering by location using the point spread function of Fermi LAT, which can be computed with \texttt{gtpsf}\footnote{\url{http://fermi.gsfc.nasa.gov/ssc/data/analysis/scitools/help/gtpsf.txt}} tool.
We use 300 energy bins from \unit[100]{MeV} to \unit[300]{GeV}, and 300 angular bins with maximal angle being $30\degree$.
We use \texttt{gtltcube}\footnote{\url{http://fermi.gsfc.nasa.gov/ssc/data/analysis/scitools/help/gtltcube.txt}} tool to compute livetime maps, which are \texttt{gtpsf} need.
Note, that we need PSFs for all Instrument Response Functions of Fermi, that is for all event classes (transient, source, clean and ultraclean) and conversion types (back and front).

@@

The exposure map can be computed with the \texttt{gtexpcube2}\footnote{\url{http://fermi.gsfc.nasa.gov/ssc/data/analysis/scitools/help/gtexpcube2.txt}} tool. We compute it with 300 energy bins from \unit[100]{MeV} to \unit[300]{GeV}.

@@

\begin{lstlisting}
gtpsf expcube=ltcube.fits outfile=psf_[IrfName].fits outtable=PSF irfs=[IrfName]
	ra=[RA] dec=[DEC] emin=100 emax=300000 nenergies=41 thetamax=30 ntheta=300
\end{lstlisting}
\texttt{[RA]} and \texttt{[DEC]} here are the coordinates of the burst, and \texttt{[IrfName]} is the Instrument Response Function, which depends on photon's event class (transient, source, clean, or ultraclean) and conversion type (back or front). The list of IRF names can be obtained with the \texttt{gtirfs}\footnote{\url{http://fermi.gsfc.nasa.gov/ssc/data/analysis/scitools/help/gtirfs.txt}} tool. We need to compute PSFs for all of them.

>>>>

Before reading the data, we need to apply the basic filtering as described in Fermi guidelines\footnote{\url{http://fermi.gsfc.nasa.gov/ssc/data/analysis/scitools/data_preparation.html}} (the actual code, and all parameters values we use to invoke Fermi Science tools are listed in appendix \ref{sec:fermiCode}).
Two tools from the Fermi Science Tools package\footnote{\url{http://fermi.gsfc.nasa.gov/ssc/data/analysis/scitools/overview.html}} are used for that task (the same filtering is applied to both burst and background data).

First, we use \texttt{gtselect}.
We filter the Earth limb emission with it below the zenith angle of $100\degree$.
We also filter out photons with transient event class for a reason described in the next section.

The second one we use is \texttt{gtmktime}. It uses the spacecraft file as well as events file, and filters out time periods when Fermi LAT was not operational. Running \texttt{gtmktime} is also required for exposure maps calculation, which we'll need later on.

Now, with basic filtering done, we can read events data from \texttt{timed.fits}. This file, however, contains observations of a large area of the sky.
Before using the data obtained from it, we should perform a more elaborate filtering by location using the point spread function of Fermi LAT, which can be computed with \texttt{gtpsf} tool.
We use 300 energy bins from \unit[100]{MeV} to \unit[300]{GeV}, and 300 angular bins with maximal angle being $30\degree$.
We use \texttt{gtltcube} tool to compute livetime maps, which are \texttt{gtpsf} need.
Note, that we need PSFs for all Instrument Response Functions of Fermi, that is for all event classes (transient, source, clean and ultraclean) and conversion types (back and front).

@@

The exposure map can be computed with the \texttt{gtexpcube2} tool. We compute it with 300 energy bins from \unit[100]{MeV} to \unit[300]{GeV}.

@@

\begin{lstlisting}
gtpsf expcube=ltcube.fits outfile=psf_[IrfName].fits outtable=PSF irfs=[IrfName]
	ra=[RA] dec=[DEC] emin=100 emax=300000 nenergies=41 thetamax=30 ntheta=300
\end{lstlisting}
\texttt{[RA]} and \texttt{[DEC]} here are the coordinates of the burst, and \texttt{[IrfName]} is the Instrument Response Function, which depends on photon's event class (transient, source, clean, or ultraclean) and conversion type (back or front). The list of IRF names can be obtained with the \texttt{gtirfs} tool. We need to compute PSFs for all of them.
